### 2016-12-11 source/discussion/figures.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Copyright (c) 2016, Perry L Miller IV
%  All rights reserved.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Discussion about figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Figures}
\label{sec:discussion_figures}

Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.

% A figure.
\begin{figure}[h]
\centering
\includegraphics{../figures/shapes.eps}
\caption{Made with LibreOffice and saved as EPS.}
\label{fig:shapes}
\end{figure}

Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

% A figure.
\begin{figure}[h]
\centering
\includegraphics[width=\linewidth,natwidth=400,natheight=300]{../figures/mountains.jpg}
\caption{Misty Mountains from Unsplash It \cite{unsplash}.}
\label{fig:mountains}
\end{figure}