### 2016-12-11 source/introduction.tex
Sorting citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Copyright (c) 2016, Perry L Miller IV
%  All rights reserved.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Introduction.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:introduction}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy
\cite{knuth-website,
knuth-fundamental-algorithms,
dirac-quantum-mechanics}.
All work and no play makes Jack a dull boy.

We're going to talk about \varA\ and \varB, which are defined elsewhere.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.