### 2016-12-11 source/appendix.tex
Added a rotated image and one that spans both columns

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Copyright (c) 2016, Perry L Miller IV
%  All rights reserved.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Appendix file.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Appendix}
\label{sec:appendix}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

% A figure using a JPG raster image file.
\begin{figure}[h]
\centering
\includegraphics[width=\linewidth,natwidth=400,natheight=300]{../figures/lake.jpg}
\caption{A Mountain Lake from Unsplash It \cite{unsplash}.}
\label{fig:lake}
\end{figure}

Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.