### 2016-12-11 source/conclusion.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Copyright (c) 2016, Perry L Miller IV
%  All rights reserved.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Conclusion file.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}
\label{sec:conclusion}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy \cite{knuth-website}.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

As noted in sections
\ref{sec:discussion_figures} and
\ref{sec:discussion_equations}
we have all work and no play making Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\subsection{Important Stuff}
\label{sec:conclusions_important_stuff}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy
\cite{einstein-electrodynamics,
knuth-website,
knuth-fundamental-algorithms,
dirac-quantum-mechanics}.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

\subsection{Really Important Stuff}
\label{sec:conclusions_really_important_stuff}

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
Figure \ref{fig:rotated_shapes} shows how we can rotate a graphic.

% A figure using an EPS vector image file.
\begin{figure}[h]
\centering
\includegraphics[width=\linewidth, angle =90]{../figures/shapes.eps}
\caption{Rotated EPS file.}
\label{fig:rotated_shapes}
\end{figure}