### 2016-12-11 source/discussion/figures.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Copyright (c) 2016, Perry L Miller IV
%  All rights reserved.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  Discussion about figures.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Figures}
\label{sec:discussion_figures}

Good, then we will shall fight in the shade.
See figures
\ref{fig:shapes},
\ref{fig:mountains}, and
\ref{fig:lake} for details.
Good, then we will shall fight in the shade.

% A figure using an EPS vector image file.
\begin{figure}[h]
\centering
\includegraphics{../figures/shapes.eps}
\caption{Made with LibreOffice and saved as EPS.}
\label{fig:shapes}
\end{figure}

Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.

% A figure using a JPG raster image file.
\begin{figure}[h]
\centering
\includegraphics[width=\linewidth,natwidth=400,natheight=300]{../figures/mountains.jpg}
\caption{Misty Mountains from Unsplash It \cite{unsplash}.}
\label{fig:mountains}
\end{figure}

Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.
Good, then we will shall fight in the shade.

All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
All work and no play makes Jack a dull boy.
Figure \ref{fig:outdoor_cafe} demonstrates how we can make an image span both columns.

% A figure that spans both columns.
\begin{figure*}[h]
\centering
\includegraphics[width=\textwidth,natwidth=800,natheight=600]{../figures/outdoor_cafe.jpg}
\caption{Outdoor Cafe from Unsplash It \cite{unsplash}.}
\label{fig:outdoor_cafe}
\end{figure*}